Star-galaxy Classification Using Deep Convolutional Neural Networks

### 2016-06-25 paper/main.tex
add final draft
added ProfessorBrunner's affiliations

\author[E. J. Kim and R. J. Brunner]{
  Edward J.~Kim$^1$\thanks{jkim575@illinois.edu} and Robert J.~Brunner$^{2}$ \\
$^1$Department of Physics, University of Illinois, Urbana, IL 61801 USA\\
$^2$Department of Astronomy, University of Illinois, Urbana, IL 61801 USA}

@@

In Figure~\ref{fig:sdss_mag}, we compare the galaxy purity
and star completeness values for ConvNet, $\rmn{TPC}_{\rmn{morph}}$, and
$\rmn{TPC}_{\rmn{phot}}$ as a
function of $r$-band magnitude for the differential counts in the SDSS data.
We note that $\rmn{TPC}_{\rmn{morph}}$ outperforms the star-galaxy
classifier used by the SDSS pipeline (\ie an object is classified as a galaxy
if $\rmn{concentration} > 0.145$) over all magnitudes.
We do not show the SDSS classifications to avoid cluttering the plots.
While ConvNet shows a similar but slightly worse performance than
$\rmn{TPC}_{\rmn{morph}}$,
the galaxy purity and star completeness values of ConvNet begin to
drop at faint magnitudes $i \la 21$.
Again, $\rmn{TPC}_{\rmn{phot}}$ performs significantly worse than both
ConvNet and TPC at bright magnitudes.
One reason that ConvNet fails to outperform $\rmn{TPC}_{\rmn{phot}}$,
especially at faint magnitudes,
might be its over-reliance on morphological features.
Near a survey's limit, the measurement uncertainties generally increase,
and morphology is not a reliable metric for star-galaxy classification.

>>>>

\author[E. J. Kim and R. J. Brunner]{
  Edward J.~Kim$^1$\thanks{jkim575@illinois.edu} and Robert J.~Brunner$^{1,2,3,4}$ \\
$^1$Department of Physics, University of Illinois, Urbana, IL 61801 USA\\
$^2$Department of Astronomy, University of Illinois, Urbana, IL 61801 USA\\
$^3$Department of Statistics, University of Illinois, Champaign, IL 61820 USA\\
$^4$National Center for Supercomputing Applications, Urbana, IL 61801 USA
}

@@

In Figure~\ref{fig:sdss_mag}, we compare the galaxy purity
and star completeness values for ConvNet, $\rmn{TPC}_{\rmn{morph}}$, and
$\rmn{TPC}_{\rmn{phot}}$ as a
function of $r$-band magnitude for the differential counts in the SDSS data.
We note that $\rmn{TPC}_{\rmn{morph}}$ outperforms the star-galaxy
classifier used by the SDSS pipeline (\ie an object is classified as a galaxy
if $\rmn{concentration} > 0.145$) over all magnitudes.
We do not show the SDSS classifications to avoid cluttering the plots.
While ConvNet shows a similar but slightly worse performance than
$\rmn{TPC}_{\rmn{morph}}$,
the galaxy purity and star completeness values of ConvNet begin to
drop at faint magnitudes $i \la 21$.
Again, $\rmn{TPC}_{\rmn{phot}}$ performs significantly worse than both
ConvNet and TPC at bright magnitudes.
One reason that ConvNet fails to outperform $\rmn{TPC}_{\rmn{phot}}$,
especially at faint magnitudes,
might be its over-reliance on morphological features.
Near a survey's limit, the measurement uncertainties generally increase,
and morphology is not a reliable metric for star-galaxy classification.
Another possibility is that data augmentation has a negative effect at faint
magnitudes, as the network may get confused by additional examples of faint
galaxies that look like point sources.
Data augmentation however is indispensable, since it improves the overall
performance greatly.
